---
output:
  pdf_document: default
  html_document:
    keep_md: yes
---
# Statistical Inference- Course Project

## The exponential distribution

(Note: we need load the `dplyr` package for data manipulation and the `matrixStats` package for the `rowSds` function.)

```{r message = FALSE, echo = FALSE}
library("dplyr")
library("matrixStats")
```

## Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponential(0.2)s.

We set the seed, the rate parameter `lambda`, the number of observations `nobs`, and the number of simulations `sims`.

```{r}
set.seed(12345); lambda = 0.2; nobs = 40; sims = 1000
```

Now we take `r options(scipen = 5); nobs*sims` random draws from the exponential distribution. We will organize this into a matrix with `r sims` rows and `r nobs` columns and view this as simulating `r sims` repetitions of a draw of `r nobs` random exponentials with rate $\lambda = `r lambda`$.

```{r}
exp <- matrix(rexp(sims*nobs, lambda), nrow = sims, ncol = nobs)
```

Now we take the mean of each row. (We'll need standard deviations and confidence interval endpoints later, so we might as well compute them now. Partway through, we convert `exp` to a data frame to be able to use the handy `mutate` function.)

```{r}
exp_mean <- rowMeans(exp)
exp_sd <- rowSds(exp)
exp <- cbind(exp, exp_mean = exp_mean, exp_sd = exp_sd)
exp <- data.frame(exp)
exp <- exp %>%
    mutate(CI_left = exp_mean - 1.96 * exp_sd / sqrt(nobs),
           CI_right = exp_mean + 1.96 * exp_sd / sqrt(nobs))
```

### 1. Show where the distribution is centered at and compare it to the theoretical center of the distribution.

The mean of our sample means is `r mean(exp$exp_mean)` versus the theoretical mean, which is
$$\frac{1}{\lambda} = \frac{1}{`r lambda`} = `r 1/lambda`.$$

### 2. Show how variable it is and compare it to the theoretical variance of the distribution.

The standard deviation of our sample means is `r sd(exp$exp_mean)` versus the theoretical standard deviation, which is
$$\frac{\sigma}{\sqrt{n}} = \frac{1/\lambda}{\sqrt{`r nobs`}} = `r (1/lambda)/sqrt(nobs)`.$$

Equivalently, the variance of our sample means is `r var(exp$exp_mean)` versus the theoretical variance, which is
$$\frac{\sigma^{2}}{n} = \frac{1/\lambda^{2}}{`r nobs`} = `r (1/lambda^2)/nobs`.$$

### 3. Show that the distribution is approximately normal.

Here is a histogram of the sample means with the normal distribution $N\left(1/\lambda, \frac{1/\lambda}{\sqrt{`r nobs`}}\right)$ superimposed.

```{r}
h <- hist(exp$exp_mean,
          main = "Histogram of sample means",
          xlab = "Mean")
xfit<-seq(min(exp$exp_mean),max(exp$exp_mean),length=100) 
yfit<-dnorm(xfit,mean = 1/lambda, sd = (1/lambda)/sqrt(nobs)) 
yfit <- yfit*diff(h$mids[1:2])*length(exp$exp_mean) 
lines(xfit, yfit, col="blue", lwd=2)
```

### 4.  Evaluate the coverage of the confidence interval for $1/\lambda$:
$$\overline{x} \pm 1.96 \frac{s}{\sqrt{n}}$$

The following code calculates the percentage of confidence intervals that capture the true (theoretical) mean.

```{r}
exp <- exp %>%
    mutate(CI_good = (CI_left < 1/lambda) & (1/lambda < CI_right))
percent_good <- mean(exp$CI_good)
```

So `r 100*percent_good`% of the simulated intervals capture the true mean $\frac{1}{\lambda} = `r 1/lambda`$.
